Comparison of Global Solution Methods to a Zero Lower Bound Model

### 2019-05-29 Paper/Paper.tex
Final edits

The \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} solution method approximates policy functions underlying the model's decision rule using an antistropic Smolyak method with Chebyshev polynomials. (See \hyperlink{Judd2}{\color{black}{Judd et al.\, 2014}} for a discussion of antistropic Smolyak methods). They approximate the exogenous state variables using Gauss-Hermite quadrature. Instead of directly computing the policy functions, they estimate functions at and away from the ZLB, which builds on \hyperlink{Christiano}{\color{black}{Christiano and Fisher (2000)}}. The idea is that the policy functions feature a kink or non-differentiability at the ZLB and regime-indexing the policy functions will yield smoother functions. This is appropriate for the choice of Chebyshev polynomial approximation, which performs better on smoother functions.

@@

\noindent
\textbf{Monetary Policy} The central bank sets the gross nominal interest rate, $i$, according to
\begin{gather}
    \label{eq:15}
    i_t=\max\{1,i_t^n\},\\
      \label{eq:16}
  i_t^n=(i^n_{t-1})^{\rho_i}(\bar{\imath}(\pi^{gap}_t)^{\phi_\pi}(y^{gdp}_{t})^{\phi_y})^{1-\rho_i}\exp(\sigma_i\varepsilon_{i,t}), 0 \leq \rho_i < 1, \varepsilon_i \sim \mathds{N}(0,1),
  \end{gather}
where $y^{gdp}$ is real GDP (i.e., output, $y$, minus the resources lost due to adjustment costs, $adj^p$), $i^n$ is the gross notional interest rate, $\bar i$ is the target values of the nominal interest rate, and $\phi_\pi$ and $\phi_y$ are the responses to the inflation and output growth gaps. A more negative net notional rate indicates that the central bank is more constrained.\\

@@

\subsection{Parameter Values} \hyperlink{Table 1}{Table 1} shows the model parameters. Parameters are from \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}}, and are chosen to be characteristic of U.S. data. The subjective discount factor, $\beta$ is set to $0.9949$, the time average of the values implied by the steady-state consumption Euler equation and the federal funds rate. The Frisch labor supply elasticity $1/\eta$ is set to 3 to match the macro estimate in \hyperlink{Peterman}{\color{black}{Peterman (2016)}}. The price elasticity of substitution $\theta$ is set to 6, matching GHLS and corresponding to to a 20\% average markup. Steady-state labor hours $\bar{n}$ is set to 1/3 of available time. The steady-state risk premium, growth rate, and inflation rate are set to the time averages of per capital real DGP growth, the percent GDP implicit price deflator, and the Baa corporate bond yield relative to the yield on the 10-Year Treasury. The capital share of income $\alpha$ is equal to the \hyperlink{Fernald}{\color{black}{Fernald (2012)}} utilization-adjusted quarterly-TFP estimates of the capital share of income from 1988-2017Q4. The other parameters are set to values consistent with posterior estimates from similar models in the literature. For the model without capital, all parameters are equivalent to those above except we set $\sigma_s=0.006$ to allow the GHLS algorithm to converge.

>>>>

The \hyperlink{Gust}{\color{black}{Gust et al.\ (2017)}} solution method approximates policy functions underlying the model's decision rule using an anisotropic Smolyak method with Chebyshev polynomials. (See \hyperlink{Judd2}{\color{black}{Judd et al.\, 2014}} for a discussion of anisotropic Smolyak methods). They approximate the exogenous state variables using Gauss-Hermite quadrature. Instead of directly computing the policy functions, they estimate functions at and away from the ZLB, which builds on \hyperlink{Christiano}{\color{black}{Christiano and Fisher (2000)}}. The idea is that the policy functions feature a kink or non-differentiability at the ZLB and regime-indexing the policy functions will yield smoother functions. This is appropriate for the choice of Chebyshev polynomial approximation, which performs better on smoother functions.

@@

\noindent
\textbf{Monetary Policy} The central bank sets the gross nominal interest rate, $i$, according to
\begin{gather}
    \label{eq:15}
    i_t=\max\{1,i_t^n\},\\
      \label{eq:16}
  i_t^n=(i^n_{t-1})^{\rho_i}(\bar{\imath}(\pi^{gap}_t)^{\phi_\pi}(y^{g}_{t})^{\phi_y})^{1-\rho_i}\exp(\sigma_i\varepsilon_{i,t}), 0 \leq \rho_i < 1, \varepsilon_i \sim \mathds{N}(0,1),
  \end{gather}
where $y^{g}$ is output growth, $i^n$ is the gross notional interest rate, $\bar{\imath}$ is the target values of the nominal interest rate, and $\phi_\pi$ and $\phi_y$ are the responses to the inflation and output growth gaps. A more negative net notional rate indicates that the central bank is more constrained.\\

@@

\subsection{Parameter Values} \hyperlink{Table 1}{Table 1} shows the model parameters. Parameters are from \hyperlink{Atkinson}{\color{black}{Atkinson et al.\ (2019)}}, and are chosen to be characteristic of U.S. data. The subjective discount factor, $\beta$ is set to $0.9949$, the time average of the values implied by the steady-state consumption Euler equation and the federal funds rate. The Frisch labor supply elasticity $1/\eta$ is set to 3 to match the macro estimate in \hyperlink{Peterman}{\color{black}{Peterman (2016)}}. The price elasticity of substitution $\theta$ is set to 6, matching GHLS and corresponding to a 20\% average markup. Steady-state labor hours $\bar{n}$ is set to 1/3 of available time. The steady-state risk premium, growth rate, and inflation rate are set to the time averages of per capital real DGP growth, the percent GDP implicit price deflator, and the Baa corporate bond yield relative to the yield on the 10-Year Treasury. The capital share of income $\alpha$ is equal to the \hyperlink{Fernald}{\color{black}{Fernald (2012)}} utilization-adjusted quarterly-TFP estimates of the capital share of income from 1988-2017Q4. The other parameters are set to values consistent with posterior estimates from similar models in the literature. For the model without capital, all parameters are equivalent to those above except we set $\sigma_s=0.006$ to allow the GHLS algorithm to converge.

### 2019-05-29 Presentation/Presentation.tex
Final edits

%\documentclass[handout, bigger]{beamer} %For Handouts
\documentclass[11pt]{beamer}
\usetheme{Singapore}
\usefonttheme[header]{structuresmallcapsserif}
\usefonttheme{professionalfonts}

@@

\begin{frame}\frametitle{Nonlinear Solution Methods}
\begin{itemize}\setlength{\itemsep}{10pt}
	\item  <1-|handout:1>As it is important to incorporate all historical data and yield acurate parameter values and predictions, the path forward seems to be in fully nonlinear models
	\item <2-|handout:1>Solving nonlinear models with projection methods (e.g. Aruoba et al., 2018; Ferna\'ndez-Villaverde et al., 2015).
	\item  <3-|handout:1>Researchers have used different methods for solving fully nonlinear models:
	\begin{enumerate} \setlength{\itemsep}{4pt}
		\item  Policy function iteration with linear interpolation (e.g., Plante et al., 2018; Richter and Throckmorton, 2016).
		\item  Alternate approximating functions: regime-indexed policy functions (Gust et al., 2017), piecewise smooth policy functions (Aruoba et al. 2018)
		\item  Alternate grid construction: Smolyak method with Chebyshev polynomials (e.g., Gust et al., 2017; Ferna\'ndez-Villaverde et al. 2015; Aruoba et al., 2018)
	\end{enumerate}
\end{itemize}
\end{frame}

@@

\end{frame}
%=========================================================================
\begin{frame}\frametitle{Atkinson et al.\ (2019)}
\begin{itemize}\setlength{\itemsep}{10pt}
\item <1-|handout:1> Time iteration with linear interpolation
\item <2-|handout:1>Directly approximates policy functions
\item <3-|handout:1> State space features evenly-spaced nodes
\item<4-|handout:1>  Exogenous variables are approximated with the Markov chain from Rouwenhorst (1995)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item Only requires interpolation along the dimensions of the endogenous state variables
\end{itemize}
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Gust et al.\ (2017)}
\begin{itemize}\setlength{\itemsep}{6pt}
\item  <1-|handout:1>Approximates policy functions using an antistropic Smolyak method with Chebyshev polynomials
\item  <2-|handout:1>Approximates exogenous state variables using Gauss-Hermite quadrature%why if we know Rouwenhorst is better?
\item  <3-|handout:1>Instead of directly computing the policy functions, they estimate functions at and away from the ZLB, building on Christiano and Fisher (2000)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item  Policy functions feature a kink or non-differentiability at the ZLB and regime-indexing the policy functions will yield smoother functions
\item Low-order Chebyshev polynomials perform better on smoother functions and are inexpensive to compute
\end{itemize}
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{This Project}
\begin{itemize}\setlength{\itemsep}{8pt}
\item  <1-|handout:1>In this project, we consider how splitting up the policy functions conditional on the ZLB impacts the speed and accuracy of the solution of a nonlinear model
\item  <2-|handout:1>We use policy function iteration on a fixed point with linear interpolation
\item  <3-|handout:1>Atkinson et al.\ (2019) provides the motivation for a single policy function;  Gust et al.\ (2017) provides the motivation for a regime-indexed policy function
\item <4-|handout:1> Solution algorithm based on Atkinson et al.\ (2019) denoted henceforth as ART; solution algorithm based on Gust et al.\ (2017) denoted henceforth as GHLS
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Our Solution Algorithm} %NEED TO PUT SOME ABOUT GHLS?
\begin{itemize}\setlength{\itemsep}{8pt}
\item   <1-|handout:1>Construct the state space with evenly-spaced nodes and approximate exogenous state variables with Rouwenhorst (1995)
\item  <2-|handout:1>Obtain initial conjectures for a set of policy functions from the log-linear solution
\item  <3-|handout:1>Using a fixed point iteration scheme, linearly interpolate and numerically integrate the policy functions each step
\begin{itemize}
\item For ART portion, we update a single policy function, and for the GHLS portion we update regime-indexed policy functions
\end{itemize}
\item <5-|handout:1> The algorithm converges once the maximum distance between successive guesses of policy functions falls below a convergence criterion

@@

\begin{frame}\frametitle{The Model (With Capital)}
\item <1-|handout:1>The households choose $\{c_t, n_t, b_t, x_t, k_t\}_{t=0}^\infty$ to maximize expected lifetime utility given by
E_0\sum_{t=0}^\infty\beta[\log(c_t-hc^a_{t-1}) - \chi n_t^{1+\eta}/(1+\eta)],
subject to their budget constraint
    c_t+x_t+b_t/(i_ts_t)=w_tn_t+r_t^kk_{t-1}+b_{t-1}/\pi_t+d_t.
\item <2-|handout:1>The nominal bond, $b$, is subject to a risk premium, $s$, that follows
  s_t = (1-\rho_s)\bar{s} + \rho_ss_{t-1} + \sigma_s\varepsilon_{s,t},
where $\bar{s}$ is the steady-state value. %An increase in $s_t$ boosts saving, which lowers period-$t$ demand.

@@

\begin{frame}\frametitle{Households}
\item <1-|handout:1>Households also face an investment adjustment cost, so the law of motion for capital is given by
  k_t = (1-\delta)k_{t-1} + x_t(1-\nu(x^g_t - 1)^2/2),\; 0 \leq \delta \leq 1.
  \end{gather*} %$x_t^g$ is investment growth relative to its steady-state and $\nu \geq 0$ scales the cost.
\item <2-|handout:1>The FOCs to each household's constrained optimization problem are
\scriptsize
  \lambda_t = c_t - hc^a_{t-1}, \\
  w_t = \chi n_t^\eta \lambda_t,\\
  1 =  \beta E_t[(\lambda_t/\lambda_{t+1})(s_ti_t/(\bar{\pi}\pi_{t+1}^{gap}))],\\
  q_t = \beta E_t[(\lambda_t/\lambda_{t+1})(r^k_{t+1}+(1-\delta)q_{t+1})],\\
  1 = q_t[1-\nu(x^g_t-1)^2/2 - \nu(x_t^g-1)x_t^g] + \nu\beta\bar{g}E_t[q_{t+1}(\lambda_t/\lambda_{t+1})(x^g_{t+1})^2(x^g_{t+1}-1)],\\
  \varphi(\pi_t^{gap}-1)\pi_t^{gap} = 1-\theta + \theta mc_t + \beta\varphi E_t[(\lambda_t/\lambda_{t+1})(\pi_{t+1}^{gap}-1)\pi_{t+1}^{gap}(y_{t+1}/y_t)].
\end{gather*} %where $1/\lambda$ is the marginal utility of consumption and $q$ is Tobin's q 
\normalsize
\begin{frame}\frametitle{Firms}
\item <1-|handout:1>The production sector consists of a continuum of monopolistically competitive intermediate goods firms and a final goods firm 
\item <2-|handout:1>Technology is $z_t = g_tz_{t-1}$, which is common across firms
\item <3-|handout:1>Deviations from the steady-state growth rate, $\bar{g}$, follow
  g_t = \bar{g} + \sigma_g\varepsilon_{g,t}.
\item <4-|handout:1>In symmetric equilibrium, the optimality conditions reduce to
\scriptsize
  y_t = (k_{t-1})^\alpha(z_tn_t)^{1-\alpha},\\
  w_t = (1-\alpha)mc_ty_t/n_t,\\
  r_t^k = \alpha mc_t y_t/k_{t-1},\\
  \varphi(\pi_t^{gap}-1)\pi_t^{gap} = 1-\theta + \theta mc_t + \beta\varphi E_t[(\lambda_t/\lambda_{t+1})(\pi_{t+1}^{gap}-1)\pi_{t+1}^{gap}(y_{t+1}/y_t)].
\normalfont
\begin{frame}\frametitle{Monetary Policy}
\begin{itemize} \setlength{\itemsep}{14pt}
\item <1-|handout:1>\ 
The central bank sets the gross nominal interest rate, $i$, according to
    i_t=\max\{1,i_t^n\},\\
  i_t^n=(i^n_{t-1})^{\rho_i}(\bar{\imath}(\pi^{gap}_t)^{\phi_\pi}(y^{gdp}_{t})^{\phi_y})^{1-\rho_i}\exp(\sigma_i\varepsilon_{i,t}). 
%where $y^{gdp}$ is real GDP (i.e., output, $y$, minus the resources lost due to adjustment costs, $adj^p$), $i^n$ is the gross notional interest rate, $\bar i$ and $\bar{\pi}$ are the target values of the inflation and nominal interest rates, and $\phi_\pi$ and $\phi_y$ are the responses to the inflation and output growth gaps. 
\item <2-|handout:1>A more negative net notional rate indicates that the central bank is more constrained

%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}\setlength{\itemindent}{-1.2em}
\begin{itemize} \setlength{\itemsep}{8pt}
\item <1-|handout:1>The aggregate resource constraint and real GDP definition are given by:
\begin{gather*}
  c_t + x_t = y_t^{gdp}\\
  y_t^{gdp} = [1 - \varphi(\pi_t^{gap}-1)^2/2]y_t
\end{gather*}
\item <2-|handout:1>The model does not have a steady-state due to the unit root in technology, $z_t$. Therefore, we define the variables with a trend in terms of technology (i.e., $\tilde{x}_t \equiv x_t/z_t$)
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}
\begin{itemize} \setlength{\itemsep}{10pt}
\item <1-|handout:1>A competitive equilibrium consists of sequences of 
\begin{enumerate} \setlength{\itemsep}{5pt}
\item quantities, $\{\tilde{c}_t, \tilde{y}_t, \tilde{y}_t^{gdp}, x^g_t, y^g_t, n_t, \tilde{k}_t, \tilde{x}_t\}_{t=0}^\infty$ \item prices, $\{\tilde{w}_t, i_t, i^n_t, \pi_t, \tilde{\lambda}_t, q_t, r^k_t, mc_t\}_{t=0}^\infty$
\item exogenous variables, $\{s_t,g_t\}_{t=0}^\infty$
\end{enumerate}
\item <2-|handout:1>that satisfy the detrended equilibrium system, given 
\begin{enumerate}\setlength{\itemsep}{5pt}
\item the initial conditions, $\{\tilde{c}_{-1}, i^n_{-1}, \tilde{k}_{-1}, \tilde{x}_{-1}, \tilde{w}_{-1}, s_0, g_0, \varepsilon_{i,0}\}$
\item three sequences of shocks, $\{\varepsilon_{g,t}, \varepsilon_{s,t}, \varepsilon_{i,t}\}_{t=1}^\infty$
\end{enumerate}
\end{itemize}
\end{frame}
%===============================================================================================================
\begin{frame}\frametitle{Parameter Values}
\begin{table}[H]
\centering
\captionsetup{justification=centering}
  \scriptsize
    \setlength{\tabcolsep}{3pt}      
  \begin{tabular}{l c c |l  c c}
    \hline
  Subjective Discount Factor & $\beta$ & $0.9949$ & Rotemberg Price Adjustment Cost & $\varphi$ & $100$ \\
   Frisch Labor Supply Elasticity & $1/\eta$ & $3$ & Inflation Gap Response & $\phi_\pi$ & $2.0$ \\
  Price Elasticity of Substitution & $\theta$ & $6$ & Output Growth Gap Response & $\phi_y$ & $0.5$ \\
  Steady-State Labor Hours & $\bar{n}$ & $1/3$ & Habit Persistence & $h$ & $0.80$ \\
  Steady-State Risk Premium & $\bar{s}$ & $1.0058$ & Risk Premium Persistence & $\rho_s$ & $0.80$ \\
  Steady-State Growth Rate & $\bar{g}$ & $1.0034$ & Notional Rate Persistence & $\rho_i$ & $0.80$ \\
  Steady-State Inflation Rate & $\bar{\pi}$ & $1.0053$ & Technology Growth Shock SD & $\sigma_g$ & $0.005$ \\
  Capital Share of Income & $\alpha$ & $0.35$ & Risk Premium Shock SD & $\sigma_s$ & $0.0085$ \\
  Capital Depreciation Rate & $\delta$ & $0.025$ & Notional Interest Rate Shock SD & $\sigma_i$ & $0.002$ \\
  Investment Adjustment Cost & $\nu$ & $4$ & & &\\
  \hline
  \end{tabular}
\normalfont
\end{table}
\begin{itemize}\setlength{\itemsep}{8pt}
\item <2-|handout:1>Parameters are from Atkinson et al.\ (2019), and are chosen to be characteristic of U.S. data
\item <3-|handout:1>For the model without capital, $\rho_s = 0.006$ %To allow the GHLS algorithm to converge
\end{itemize}
\end{frame}
%===================================================================================
\begin{frame}\frametitle{Discretized State Space (No Capital)}

@@

%===============================================================================================================
\begin{frame}\frametitle{Solving the Model}

\begin{enumerate}\setlength{\itemsep}{10pt}
  \item <1-|handout:1>Calculate $z(s')$ for each realization of the state
  \item <2-|handout:1>Find the updated policy function, $n(s')$, for each state
  \item <3-|handout:1>Numerical Integration:
  \begin{itemize}\setlength{\itemsep}{5pt}
    \item First integrate across the continuous random variable, $z$, conditional on the future realizations of the discrete stochastic variable, $s'$.%, to obtain:
        %\begin{gather*}
         % E[\Phi(z')|s'=1], \quad E[\Phi(z')|s'=2]
        %\end{gather*}
    \item Then weight the conditional expectations by their corresponding likelihood. %The conditional expectations are:
  %  \begin{gather*}
    %  E\left[\Phi(z')|s = i\right] = p_{i1}E[\Phi(z')|s'=1] + p_{i2}E[\Phi(z')|s'=2]
   % \end{gather*}
  \end{itemize}
\end{enumerate}

@@

\begin{itemize}\setlength{\itemsep}{8pt}
  \item<1-|handout:1> Used to approximate an exogenous $AR(1)$ process
  \item <2-|handout:1>Kopecky and Suen (2010) show the Rouwenhorst method outperforms other approximations of an $AR(1)$ process
  \item <3-|handout:1>The approximation is a Markov switching process like the time-varying intercept example, but with $n$ states
  \item <4-|handout:1>The method determines the bounds of the exogenous state variables, the nodes, and the transition probabilities
  \item <5-|handout:1>Let $z\sim AR(1)$ with persistence $\rho$, mean $\mu_z$, and variance %Idea for approximation of an AR(1) process
  \begin{gather*}
    \sigma_z^2 = \sigma_\varepsilon^2/(1-\rho^2).
  \end{gather*}
  \item   <6-|handout:1>The $n$ states for the discretized process $z$ are evenly spaced on $[\mu_z-\sigma_z\sqrt{n-1},\mu_z+\sigma_z\sqrt{n-1}]$
\end{itemize}

>>>>

%\documentclass[handout, 11pt]{beamer} %For Handouts
\documentclass[11pt]{beamer}
\usetheme{Singapore}
\usefonttheme[header]{structuresmallcapsserif}
\usefonttheme{professionalfonts}

@@

%=========================================================================
\begin{frame}\frametitle{The Model (With Capital)}
\begin{itemize}\setlength{\itemsep}{6pt}
\item <1-|handout:1>The households choose $\{c_t, n_t, b_t, x_t, k_t\}_{t=0}^\infty$ to maximize expected lifetime utility given by
\begin{gather*}
E_0\sum_{t=0}^\infty\beta[\log(c_t-hc^a_{t-1}) - \chi n_t^{1+\eta}/(1+\eta)],
\end{gather*} 
subject to their budget constraint
\begin{gather*}
    c_t+x_t+b_t/(i_ts_t)=w_tn_t+r_t^kk_{t-1}+b_{t-1}/\pi_t+d_t.
  \end{gather*} 
\item <2-|handout:1>The nominal bond, $b$, is subject to a risk premium, $s$, that follows
\begin{gather*}
  s_t = (1-\rho_s)\bar{s} + \rho_ss_{t-1} + \sigma_s\varepsilon_{s,t},
\end{gather*}
where $\bar{s}$ is the steady-state value. %An increase in $s_t$ boosts saving, which lowers period-$t$ demand.
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Households}
\begin{itemize}\setlength{\itemsep}{8pt}
\item <1-|handout:1>Households also face an investment adjustment cost, so the law of motion for capital is given by
\begin{gather*}
  k_t = (1-\delta)k_{t-1} + x_t(1-\nu(x^g_t - 1)^2/2),\; 0 \leq \delta \leq 1.
  \end{gather*} %$x_t^g$ is investment growth relative to its steady-state and $\nu \geq 0$ scales the cost.
\item <2-|handout:1>The FOCs to the representative household's constrained optimization problem are
\scriptsize
\begin{gather*}
  \lambda_t = c_t - hc^a_{t-1}, \\
  w_t = \chi n_t^\eta \lambda_t,\\
  1 =  \beta E_t[(\lambda_t/\lambda_{t+1})(s_ti_t/(\bar{\pi}\pi_{t+1}^{gap}))],\\
  q_t = \beta E_t[(\lambda_t/\lambda_{t+1})(r^k_{t+1}+(1-\delta)q_{t+1})],\\
  1 = q_t[1-\nu(x^g_t-1)^2/2 - \nu(x_t^g-1)x_t^g] + \nu\beta\bar{g}E_t[q_{t+1}(\lambda_t/\lambda_{t+1})(x^g_{t+1})^2(x^g_{t+1}-1)].\\
\end{gather*} %where $1/\lambda$ is the marginal utility of consumption and $q$ is Tobin's q 
\normalsize
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Firms}
\begin{itemize}\setlength{\itemsep}{8pt}
\item <1-|handout:1>The production sector consists of a continuum of monopolistically competitive intermediate goods firms and a final goods firm 
\item <2-|handout:1>Technology is $z_t = g_tz_{t-1}$, which is common across firms
\item <3-|handout:1>Deviations from the steady-state growth rate, $\bar{g}$, follow
\begin{gather*}
  g_t = \bar{g} + \sigma_g\varepsilon_{g,t}.
\end{gather*}
\item <4-|handout:1>In symmetric equilibrium, the optimality conditions reduce to
\scriptsize
\begin{gather*}
  y_t = (k_{t-1})^\alpha(z_tn_t)^{1-\alpha},\\
  w_t = (1-\alpha)mc_ty_t/n_t,\\
  r_t^k = \alpha mc_t y_t/k_{t-1},\\
  \varphi(\pi_t^{gap}-1)\pi_t^{gap} = 1-\theta + \theta mc_t + \beta\varphi E_t[(\lambda_t/\lambda_{t+1})(\pi_{t+1}^{gap}-1)\pi_{t+1}^{gap}(y_{t+1}/y_t)].
\end{gather*}
\normalfont
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Monetary Policy}
\begin{itemize} \setlength{\itemsep}{14pt}
\item <1-|handout:1>\ 
The central bank sets the gross nominal interest rate, $i$, according to
\begin{gather*}
    i_t=\max\{1,i_t^n\},\\
  i_t^n=(i^n_{t-1})^{\rho_i}(\bar{\imath}(\pi^{gap}_t)^{\phi_\pi}(y^{g}_{t})^{\phi_y})^{1-\rho_i}\exp(\sigma_i\varepsilon_{i,t}). 
\end{gather*}
%where $y^{gdp}$ is real GDP (i.e., output, $y$, minus the resources lost due to adjustment costs, $adj^p$), $i^n$ is the gross notional interest rate, $\bar i$ and $\bar{\pi}$ are the target values of the inflation and nominal interest rates, and $\phi_\pi$ and $\phi_y$ are the responses to the inflation and output growth gaps. 
\item <2-|handout:1>A more negative net notional rate indicates that the central bank is more constrained
\end{itemize}
\end{frame}

%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}\setlength{\itemindent}{-1.2em}
\begin{itemize} \setlength{\itemsep}{8pt}
\item <1-|handout:1>The aggregate resource constraint and real GDP definition are given by:
\begin{gather*}
  c_t + x_t = y_t^{gdp}\\
  y_t^{gdp} = [1 - \varphi(\pi_t^{gap}-1)^2/2]y_t
\end{gather*}
\item <2-|handout:1>The model does not have a steady-state due to the unit root in technology, $z_t$. Therefore, we define the variables with a trend in terms of technology (i.e., $\tilde{x}_t \equiv x_t/z_t$)
\end{itemize}
\end{frame}
%=========================================================================
\begin{frame}\frametitle{Competitive Equilibrium}
\begin{itemize} \setlength{\itemsep}{10pt}
\item <1-|handout:1>A competitive equilibrium consists of sequences of 
\begin{enumerate} \setlength{\itemsep}{5pt}
\item quantities, $\{\tilde{c}_t, \tilde{y}_t, \tilde{y}_t^{gdp}, x^g_t, y^g_t, n_t, \tilde{k}_t, \tilde{x}_t\}_{t=0}^\infty$ \item prices, $\{\tilde{w}_t, i_t, i^n_t, \pi_t, \tilde{\lambda}_t, q_t, r^k_t, mc_t\}_{t=0}^\infty$
\item exogenous variables, $\{s_t,g_t\}_{t=0}^\infty$
\end{enumerate}
\item <2-|handout:1>that satisfy the detrended equilibrium system, given 
\begin{enumerate}\setlength{\itemsep}{5pt}
\item the initial conditions, $\{\tilde{c}_{-1}, i^n_{-1}, \tilde{k}_{-1}, \tilde{x}_{-1}, \tilde{w}_{-1}, s_0, g_0, \varepsilon_{i,0}\}$
\item three sequences of shocks, $\{\varepsilon_{g,t}, \varepsilon_{s,t}, \varepsilon_{i,t}\}_{t=1}^\infty$
\end{enumerate}
\end{itemize}
\end{frame}
%===============================================================================================================
\begin{frame}\frametitle{Parameter Values}
\begin{table}[H]
\centering
\captionsetup{justification=centering}
  \scriptsize
    \setlength{\tabcolsep}{3pt}      
  \begin{tabular}{l c c |l  c c}
    \hline
  Subjective Discount Factor & $\beta$ & $0.9949$ & Rotemberg Price Adjustment Cost & $\varphi$ & $100$ \\
   Frisch Labor Supply Elasticity & $1/\eta$ & $3$ & Inflation Gap Response & $\phi_\pi$ & $2.0$ \\
  Price Elasticity of Substitution & $\theta$ & $6$ & Output Growth Gap Response & $\phi_y$ & $0.5$ \\
  Steady-State Labor Hours & $\bar{n}$ & $1/3$ & Habit Persistence & $h$ & $0.80$ \\
  Steady-State Risk Premium & $\bar{s}$ & $1.0058$ & Risk Premium Persistence & $\rho_s$ & $0.80$ \\
  Steady-State Growth Rate & $\bar{g}$ & $1.0034$ & Notional Rate Persistence & $\rho_i$ & $0.80$ \\
  Steady-State Inflation Rate & $\bar{\pi}$ & $1.0053$ & Technology Growth Shock SD & $\sigma_g$ & $0.005$ \\
  Capital Share of Income & $\alpha$ & $0.35$ & Risk Premium Shock SD & $\sigma_s$ & $0.0085$ \\
  Capital Depreciation Rate & $\delta$ & $0.025$ & Notional Interest Rate Shock SD & $\sigma_i$ & $0.002$ \\
  Investment Adjustment Cost & $\nu$ & $4$ & & &\\
  \hline
  \end{tabular}
\normalfont
\end{table}
\begin{itemize}\setlength{\itemsep}{8pt}
\item <2-|handout:1>Parameters are from Atkinson et al.\ (2019), and are chosen to be characteristic of U.S. data
\item <3-|handout:1>For the model without capital, $\sigma_s = 0.006$ %To allow the GHLS algorithm to converge
\end{itemize}
\end{frame}
%===============================================================================================================

\begin{frame}\frametitle{Nonlinear Solution Methods}
\begin{itemize}\setlength{\itemsep}{10pt}
	\item  <1-|handout:1>As it is important to incorporate all historic data and yield accurate parameter values and predictions, the path forward seems to be in fully nonlinear models
	\item <2-|handout:1>Solving nonlinear models with projection methods:
\begin{enumerate}
		\item  <3-|handout:1>Policy function iteration with linear interpolation on an evenly spaced grid (e.g., Plante et al., 2018; Richter and Throckmorton, 2016).
		\item  <3-|handout:1>Alternate approximating functions: regime-indexed policy functions (Gust et al., 2017), piecewise smooth policy functions (Aruoba et al. 2018)
		\item  <3-|handout:1>Alternate grid construction and policy function evaluation method: Smolyak method with Chebyshev polynomials (e.g., Gust et al., 2017; Ferna\'ndez-Villaverde et al. 2015; Aruoba et al., 2018)
	\end{enumerate}
\end{itemize}
\end{frame}

@@

\begin{frame}\frametitle{Atkinson et al.\ (2019)}
\begin{itemize}\setlength{\itemsep}{10pt}
\item <1-|handout:1> Time iteration with linear interpolation
\item <2-|handout:1>Directly approximates policy functions
\item <3-|handout:1> State space features evenly-spaced nodes
\item<4-|handout:1>  Exogenous variables are approximated with the Markov chain from Rouwenhorst (1995)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item Only requires interpolation along the dimensions of the endogenous state variables
\begin{frame}\frametitle{Gust et al.\ (2017)}
\item  <1-|handout:1>Approximates policy functions using an anisotropic Smolyak method with Chebyshev polynomials
\item  <2-|handout:1>Approximates exogenous state variables using Gauss-Hermite quadrature%why if we know Rouwenhorst is better?
\item  <3-|handout:1>Instead of directly computing the policy functions, they estimate functions at and away from the ZLB, building on Christiano and Fisher (2000)
  \begin{itemize}\setlength{\itemsep}{4pt}
\item  Policy functions feature a kink or non-differentiability at the ZLB and regime-indexing the policy functions will yield smoother functions
\item Low-order Chebyshev polynomials perform better on smoother functions and are inexpensive to compute
\begin{frame}\frametitle{This Project}
\item  <1-|handout:1>In this project, we consider how splitting up the policy functions conditional on the ZLB impacts the speed and accuracy of the solution of a nonlinear model
\item  <2-|handout:1>We use policy function iteration on a fixed point with linear interpolation
\item  <3-|handout:1>Atkinson et al.\ (2019) directly computes the policy functions;  Gust et al.\ (2017) indirectly computes the policy functions through regime-indexed policy functions
\item <4-|handout:1> Solution algorithm based on Atkinson et al.\ (2019) denoted henceforth as ART; solution algorithm based on Gust et al.\ (2017) denoted henceforth as GHLS

@@

\end{frame}%=========================================================================
\begin{frame}\frametitle{Our Solution Algorithm} %NEED TO PUT SOME ABOUT GHLS?
\item   <1-|handout:1>Construct the state space with evenly-spaced nodes and approximate exogenous state variables with Rouwenhorst (1995)
\item  <2-|handout:1>Obtain initial conjectures for a set of policy functions from the log-linear solution
\item  <3-|handout:1>Using a fixed point iteration scheme, linearly interpolate and numerically integrate the policy functions each step
\begin{itemize}
\item For ART portion, we update a single policy function, and for the GHLS portion we update regime-indexed policy functions

@@

\item <5-|handout:1> The algorithm converges once the maximum distance between successive guesses of policy functions falls below a convergence criterion

%===================================================================================
\begin{frame}\frametitle{Discretized State Space (No Capital)}

@@

@@

\begin{itemize}\setlength{\itemsep}{12pt}
  \item<1-|handout:1> Used to approximate an exogenous $AR(p)$ process
  \item <2-|handout:1>Kopecky and Suen (2010) show the Rouwenhorst method outperforms other approximations of an $AR(1)$ process
  \item <3-|handout:1>The method determines the bounds of the exogenous state variables, the nodes, and the transition probabilities
\end{itemize}
\end{frame}

%===============================================================================================================
%===============================================================================================================

\begin{frame}\frametitle{Example: Time Iteration}

On iteration $q$, solve for the ${\cal P}_c^q(g,s,mp,i^n)$ that \textit{satisfies} equilibrium
\begin{enumerate}\setlength{\itemsep}{12pt}
  \item<1-|handout:1> Use log-linear solution on each node to obtain ${\cal P}_c^0$
  \begin{itemize}
    \item Local: ${\cal P}_c^1={\cal P}_c^0$
    \item Global: set initial basis $\hat{\eta}^0$%$\hat{\eta}^0 = ({X^T}X)^{-1}{X^T}\operatorname{vec}({\cal P}_c^0)$ so ${\cal P}_c^1=X\hat{\eta}^0$
  \end{itemize}
  \item<2-|handout:1> Solve for $g'$, $s'$, $mp'$, ${i^n}'$, given $\varepsilon_g'$, $\varepsilon_s'$, $\varepsilon_i'$ \label{updated vars}
  \item<3-|handout:1> Find ${\cal P}_c^q(g',s',mp',{i^n}')$ given the updated state
  \begin{itemize}
    \item Local: use piecewise linear interpolation
    \item Global: update the basis %so ${\cal P}_c^q(g',s',mp',{i^n}') = X'\hat{\eta}^{q-1}$
  \end{itemize}
  \item<4-|handout:1> Evaluate expectations (Rouwenhorst or Gauss Hermite) \label{expectation}
    %\begin{gather*}
      %E[\Phi(z')] = \beta E[{\cal P}_c^q(k',z')^{-\sigma}(\alpha z'k'^{\alpha-1}+1-\delta)]
    %\end{gather*}
\end{enumerate}

\end{frame}
\setcounter{savedenumi}{\value{enumi}}
